---
title: "Analyse wolves data models fitted to packs"
output: html_notebook
---

```{r setup}

library(tidyverse)
library(magrittr)

# A basic population variance function
# NB this contrasts with R's default cov() which uses sample variance.
popVar <- function(x){sum((x - mean(x)) ^2) / length(x)}

popCov <- function(x, y=NULL) {
   cov(x,y)*(nrow(x)-1)/nrow(x)
}

```


## Load the previous runs

```{r}

# load the source data by region
raw_source_data <- read.csv(file = "wolves/wolves_sources.csv", header = TRUE)

# load the full raw data
raw_wolves_consumer_all_data <- read.csv(file = "wolves/wolves_consumer.csv", header = TRUE)

load(file = "wolves-normal-process-only-region-1.RData")

load(file = "wolves-normal-process-only-region-2.RData")

load(file = "wolves-normal-process-only-region-3.RData")
```


### Analyze diagnostics and output

Set the options for the output which is common to all subsequent calls `output_stats()`.

```{r, eval=TRUE}
output_options <- list(summary_save = FALSE,
                       summary_name = "summary_statistics",
                       sup_post = FALSE,
                       plot_post_save_pdf = TRUE,
                       plot_post_name = "posterior_density",
                       sup_pairs = FALSE,
                       plot_pairs_save_pdf = TRUE,
                       plot_pairs_name = "pairs_plot",
                       sup_xy = TRUE,
                       plot_xy_save_pdf = FALSE,
                       plot_xy_name = "xy_plot",
                       gelman = TRUE,
                       heidel = FALSE,
                       geweke = TRUE,
                       diag_save = TRUE,
                       diag_name = "diagnostics",
                       indiv_effect = FALSE,
                       plot_post_save_png = FALSE,
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE,
                       diag_save_ggmcmc = FALSE, 
                       return_obj = TRUE)
```

Then you can call `output_JAGS` to process diagnostics, summary statistics, and create posterior density plots:
```{r, eval=TRUE}
df.stats_region_1 <- output_stats(jags_region_1, 
                                  mix_1, 
                                  source_1, 
                                  output_options)

df.stats_region_2 <- output_stats(jags_region_2, 
                                  mix_2, 
                                  source_2, 
                                  output_options)

df.stats_region_3 <- output_stats(jags_region_3, 
                                  mix_3, 
                                  source_3, 
                                  output_options)



```

## try to calculate the summed variances of the proportions by pack

```{r}

wolves_summary <- bind_rows(
  as.data.frame(df.stats_region_1),
  as.data.frame(df.stats_region_2),
  as.data.frame(df.stats_region_3))

wolves_summary <- as_tibble(wolves_summary,
                            rownames = "post")

wolves_packs <- filter(wolves_summary, 
                       grepl('p.Individual', post))

indiv_labels <- str_extract_all(wolves_packs$post, 
                    "\\d", 
                    simplify = T)

# combine the numeric strings by row and coerce to numberic
wolves_packs %<>% mutate(Individual = 
                           as.numeric(
                             apply(indiv_labels, 
                                   1, 
                                   paste0, 
                                   collapse = ""))) 

```

Merge with the pack data so we can group it an calculate pack-level variance

```{r}

wolves_packs <- left_join(wolves_packs,
                          select(raw_wolves_consumer_all_data,
                                 Individual, 
                                 Pack))

```

And now calculate variance in the diet proportions

```{r}

# reshape the mean proportions into a nx3 matrix of 
# individuals across the 3 sources and calculate the 
# covariance matrix and return its trace.

pack_var_p <- wolves_packs %>% 
  group_by(Pack) %>% 
  summarise(var_p = sum(diag(cov(matrix(Mean, ncol = 3)))))

```

Calculate pack level variance from the raw isotope data

```{r}

pack_var_y <- raw_wolves_consumer_all_data %>% 
  group_by(Pack) %>%
  summarise(var_y = sum(diag(popCov(cbind(d13C, d15N)))))

```

Join the proportion variance and raw isotope variance estimates

```{r}

# join the dataframes containing the variance in 
# proportions with the one containing variance in 
# raw consumer isotope data. 
pack_var <- inner_join(pack_var_p, pack_var_y, 
                       by = "Pack")

# extract the pack-to-region key to allow a lookup
pack_region_key <- raw_wolves_consumer_all_data %>%
  group_by(Pack) %>%
  summarise(Pack = Pack[1], Region = Region[1])

# add the region data to the consumer variance dataset
pack_var <- left_join(pack_var, 
                       pack_region_key, 
                       by = "Pack")
```

Calculate the source isotope variance

```{r}

var_S <- raw_source_data %>% group_by(Region) %>% 
  summarise(var_S = popVar(Meand13C) + sum(SDd13C^2) +
              popVar(Meand15N) + sum(SDd15N^2))

# Add source variance to the consumer variance dataset
pack_var <- left_join(pack_var, var_S, 
                       by = "Region")

```


And plot it

```{r}

g1 <- ggplot(data = pack_var,
             mapping = aes(x = var_p, 
                           y = var_y)) + 
  geom_point() + 
  geom_smooth(method = "lm")

print(g1)

```

Correct by source variation

```{r}

g2 <- ggplot(data = pack_var,
             mapping = aes(x = (3/2) * var_p,
                           y = var_y / var_S)
             ) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  geom_abline(slope = 1, intercept = 0)

print(g2)



```





```{r previous-code, eval = F}

pack_summaries <- wolves_packs %>% group_by(pack) %>% 
  summarise(sum(diag(cov(matrix(Mean, ncol = 34)))))

pack_iso_var <- mix$data %>% group_by(Pack, ) %>% 
  summarise(iso_var = popVar(d13C) + popVar(d15N), 
            region = Region[1])

# get the baseline variation by region
source_variation_region_1 <- source_1$S_MU %>% as_tibble %>%
  summarise(source_var = popVar(Meand13C) + popVar(Meand15N))


```


