---
title: "Simluate over Number of Sources"
output: html_notebook
---

```{r setup}

# these packages are required but do not need to be loaded.
# If you do load them, MCMCpack has to be loaded before tidyverse
# else it masks an important function select()
# install.packages(c("SIBER", "MCMCpack"))
# library(SIBER)
# library(MCMCpack)

# these packages are required and need to be loaded.
library(tidyverse)
library(magrittr)
library(patchwork)

# pretty summary table printing these need to be installed, but not loaded
# install.packages("sjplot) # is sufficient to pull the others
# library(sjPlot)
# library(sjmisc)
# library(sjlabelled)




# source the R files while this is not a compiled package
# source("../R/simPopEquilibrium.R")

```


A simulation of populations of consumers over a range of food sources.

There are basically two ways to do this:

1. Generate a single set of diets and vary the variation among food sources; and repeat for a number of diets of identical variation. This will produce a nested structure with the ethos similar to the Hoeinghaus paper example.
2. Generate a new set of diets each time. This option will result in more noise in the system.

In this example, I generate only a single set of diet proportions and vary the sources over an increasing range (corresponds to variance). The example below is for a single isotope tracer only: a little more thought is required for higher number of tracers.



Define the properties of the consumers

```{r}

# how many populations to simulate
n_pops <- 500

# how many consumers per population
n_individuals <- 50

# number of food sources they eat from
K <- ceiling(runif(n_pops, min = 2, max = 20 ))

# set how many isotopes are used in each population and create the tibble
iso_tbl <- tibble(n_individuals, K, n_iso = rep(2, n_pops) )

```

## Generate the dietary proportions

```{r}

# K is a vector 1:3
# the alphas are stored in list aa, which can be of unequal length

aa <- 3

generateProportions <- function(n, K, aa){
  
  alpha_diet <- rep(aa, K)
  
  return(MCMCpack::rdirichlet(n = n, alpha = alpha_diet))
  
}

# use cross_df to set up all combinations, but we have to generate an index
# for the list aa and later insert the corresponding values from aa in 
# the tibble and remove the idx column.
# aj <- cross_df(list(idx = 1:4, K = 1:3))
# bb <- aj %>% mutate(alphas = aa[idx]) %>% select(-idx) 

iso_tbl %<>% mutate(p = pmap( list(n = n_individuals, K = K, aa = aa), generateProportions))



```



## Generate the Source isotope values


```{r}

# the max range of the isotopes. Here i am rescaling real isotope values so that
# their S_min is zero. I am imaging these to be d13C and d15N in this example.
# I generate  range of them from 0 to 20 for d13C and 0 to 10 for d15N.


# a wrapper function to pull runifs for the range of source isotoeps
genSourceRange <- function(n_iso, max_val = 30){
  runif(n_iso, min = 1, max = max_val)
  }

# generate maximum source isotope values from which to draw Sources in the  
# subsequent step.
iso_tbl %<>% mutate(S_max = 10)


# a function to generate the sources by drawing them from runif()
generateSources <- function(K = 3, S_max = c(1, 10), setSeed = NULL, ...){
  
  if(!is.null(setSeed)) {set.seed(setSeed)}
  
  # generate a matrix of runif values
  # with K rows and length(S) columns.
  # Each column i has runif max = S_max[i]
  matrix(runif(n = K * length(S_max), min = 0, max = S_max),
         ncol = length(S_max), byrow = TRUE)
  # print(c(K,S_max))
  
}


# ------------------------------------------------------------------------------

# use pmap to loop over the rows of our tibble, select the K and S_max
# columns and use those to draw a set of runif distributed isotope values for
# each of the K sources and number of isotopes (length of S_max)
iso_tbl %<>% mutate(S = pmap( list(K, S_max), generateSources, setSeed = NULL))



```


## Calculate consumer isotope values

A function to generate the consumer isotope values

```{r}
simPopEquilibrium <- function(p = MCMCpack::rdirichlet(1, rep(1, 3)),
                              S = matrix(runif(6, min = -10, max = 0),
                                         nrow = 3, ncol = 2), TDF = 0, ...) {

  

  # Calculate the Stable isotope values of the consumers as a
  # p weighted sum of food source isotoep values (S) after applying the
  # Trophic Discrimination factor (TDF)
  consumerIso <- p %*% (S + TDF)

  return(consumerIso)
}

```


```{r}

# call the function simPopEquilibrium which we sourced from "R/" directory
# in the setup chunk. This function calculates the consumer isotope values.
# This function is mapped over each row in our dataset.
iso_tbl %<>% mutate(X = pmap(list(p, S), simPopEquilibrium))


```


# Calculate Variances

```{r dirichlet-moments}

# given a vector of alphas calculate the first moments
dirichletMomentsAlphas <- function(aa) {
  
  a_0 <- sum(aa)
  
  # the mean
  mu_X <- aa / a_0
  
  # variance
  var_X <- aa * (aa - a_0) / (a_0^2 * (a_0 + 1))
  
  # covariance
  a_ij <- aa %*% t(aa)
  diag(a_ij) <- 0 # set i=j to zero
  
  cov_X <- -a_ij / (a_0^2 * (a_0 + 1))
  
  return(list(mu_X, var_X, cov_X))
  
  
}
  

dirichletMomentsAlphas(rep(3, 5))
                        
```


There are numerous ways in which variance can be calculated on these multivariate data.

```{r}

# The Trace (sum of the diagonal) of the covariance matrix of the 
# dietary proportions. Which will approximate the sample estimate
iso_tbl %<>% 
  mutate(p_trace = pmap(list(p), ~unlist(sum(diag(cov(.))))) %>% unlist)

# calculate the area of the Standard Ellipse Area of the isotope data
# NB only works in two dimensions for SIBER but it can easily be 
# extended to Z dimensions. We use an internal if() statement to check 
# before calculating the area, else it returns NA.
doSEA <- function(x) {
  if(ncol(x) == 2) {
    out <- SIBER::sigmaSEA(cov(x))$SEA
  
  } else {
  out <- NA
  }
  return(out)
}

# map this function over all rows of our tibble
iso_tbl %<>% 
  mutate(SEA = pmap(list(X), ~doSEA(.)) %>% unlist,
         X_var = pmap(list(X), ~sum(cov(.))) %>% unlist)


# calculate sum of the covariance matrix of the food sources mapped over
# each row.
iso_tbl %<>% 
  mutate(S_var = pmap(list(S), ~sum(cov(.))) %>% unlist)

```


## Plot effect of increasing source variation on consumer isotopes

### Standard Ellipse Area of consumer Vs SEA


```{r}

g1 <- ggplot(data = iso_tbl, 
             mapping = aes(x = K, y = SEA)) + 
  geom_point()

print(g1)

```

### Trace of Consumer Vs Trace of Sources

Plotted on log-log scale since our expectation from Jeff's model is that the relationship would be multiplicative. $\sigma^2_X \propto a\sigma^2_S$ where $a$ is a scaling constant that accounts for the other sources of variation. Furthermore, our expectation would be that of isometric scaling so that in $\sigma^2_X \propto a(\sigma^2_S)^b$ the scaling term $b = 1$. This can be formally tested by linear regression on the log-log scale $\log(\sigma^2_X) = a + b\ln(\sigma^2_S)$, where the slope of $\sigma^2_S$ is expected to be 1.
 
```{r}

K_line <- tibble(K = 1:max(iso_tbl$K))
K_line %<>% mutate(X_var = (K) / (K-1))


g2 <- ggplot(data = iso_tbl, 
             mapping = aes(x = K, y = X_var)) + 
  geom_point() + 
  geom_line(data = K_line, 
            mapping = aes(x = K, 
                          y = X_var)) + 
  scale_x_log10() + 
  scale_y_log10()

print(g2)

ggsave(g2, filename = "X_var-by-K.eps")

# and a linear model
m2 <- glm(log(X_var) ~ log(I(K / (K - 1))), data = iso_tbl)

# the sjPlot::tab_model() function is not outputting to Rnotebook
# m2 %>% sjPlot::tab_model()

summary(m2)

```


### Correct for N Sources (K)

It is clear from this example, generated with a common set of dietary proportions across all simulations, that dividing by the source isotope variance $\sigma^2_S$ removes the correlation.

```{r}

g3 <- ggplot(data = iso_tbl, 
             mapping = aes(x = K, y = X_var *  (K-1) / (K))) + 
  geom_point() + 
  # geom_smooth(method = "lm") + 
  ggtitle("Raw scale")

# and the same plot on log scale
g4 <- g3 + scale_x_log10() + 
  scale_y_log10() + 
  ggtitle("Log scale")

  
# plot side by side  
(g3 | g4)


```






